---
title: "R Notebook"
output: html_notebook
---

# Overview


# Preprocessing

## Packages

Load the necessary packages.
```{r, message=FALSE}
library(tidyverse)
```

## Data Path

Establish file path to the project directory. This will be used to access data.
```{r}
project.path <- rprojroot::find_rstudio_root_file()
data.path <- file.path(project.path, "data")
```

## Functions

General data preprocessing function.
```{r}
prep <- function(x) {
  x %>% 
  rename_all(tolower) %>% 
  mutate_if(is.character,
            list(~tolower(.) %>% trimws)) %>% 
  mutate_if(is.character,
            list(~str_replace_all(., " ", "_"))) 
}
```

## Import Dat

### Master Taxonomic Table

Import the "master" taxonomic table. This table includes all of the macroinvertebrate taxonomic hierarchy and attributes.
```{r}
master.df <- file.path(data.path,
                       "macro_taxa-hier-attributes_2019-04-10.csv") %>% 
  read.csv(stringsAsFactors = FALSE,
           na.strings=c(""," ","NA")) %>% 
  prep()
```

### Abundance Data

Import the the observed taxonomic abundances.
```{r}
abund.df <- file.path(data.path,
                       "wappinger-2017_species-raw.csv") %>% 
  read.csv(stringsAsFactors = FALSE,
           na.strings=c(""," ","NA")) %>% 
  prep()
```


## Join Tables

```{r}
join.df <- abund.df %>% 
  left_join(master.df, by = c("macro_genspecies"))
```

Check to see if any of the taxa did not merge correctly with the master taxonomic table. 
```{r}
check.vec <- join.df %>% 
  filter(is.na(kingdom)) %>% 
  distinct(macro_genspecies) %>% 
  pull()

if (length(check.vec) == 0) {
  "All taxa merged correctly."
} else {
  paste("The following taxa are missing fromt he Master Taxonomic Table:",
        paste(check.vec,
              sep = ", ",
              collapse = ", "))
}
```

Finalize the data frame before exporting.
```{r}
final.df <- join.df %>% 
  unite(station, c("basin", "location", "rivmile")) %>% 
  select(station, order, family, macro_genspecies, indiv) %>% 
  spread(station, indiv, fill = 0) %>% 
  mutate_at(c("order", "family", "macro_genspecies"),
            ~str_to_sentence(.) %>% 
              str_replace_all("_", " ")) %>% 
  rename(Order = order,
         Family = family,
         Taxon = macro_genspecies)
```

Export `final.df` as a CSV.
```{r}
data.table::fwrite(final.df,
                   file.path(data.path,
                            "output",
                            paste0("wappinger-2017_taxa_", Sys.Date(), ".csv")),
                  )
```

